---
title: "Econ302 - R Assignment - 1"
author: "Mert Yazal"
date: "10/24/2019"
output: 
  rmarkdown::html_document:
    theme: lumen
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, echo=FALSE,warning=FALSE, message=FALSE}
library(AER)
library(data.table)
library(ggplot2)
library(readxl)
library(knitr)
library(QuantPsyc)
library(rgl)
library(rglwidget)
```

```{r librariess, echo=FALSE,warning=FALSE, message=FALSE}
lagpad <- function(x, k) {
  if (k>0) {
    return (c(rep(NA, k), x)[1 : length(x)] );
  }
  else {
    return (c(x[(-k+1) : length(x)], rep(NA, -k)));
  }
}
#function for adding lags
```

```{r load myData, include=FALSE}
data <- read_xlsx("r_assignment.xlsx")
model1 <- lm(Y ~ X + Z, data = data)
fit <- lm(Y ~ X, data = data)
fit2 <- lm( Y ~ Z, data = data)
fit <- lm(data$Y ~ data$X, data = data) 
data$predicted <- predict(fit)   
data$residuals <- residuals(fit) 
data$predicted <- predict(fit2)  
data$residuals <- residuals(fit2) 
```

# Task 1
Task 1 requires OLS Estimation, DW and Breusch-Pagan Test.


### OLS Estimation {.tabset}

Let's estimate our regression model $Y_t = \beta_0 + \beta_1{X_t} + \beta_2{Z_t} + u_t$, and plot what the series are revealing:

#### OLS

```{r ols555, echo=FALSE, warning=FALSE, message=FALSE}
summary(model1)
```

We obtain: $\widehat{Y_{t}} = \widehat{\beta_0} + \widehat{\beta_1}X_{t} + \widehat{\beta_2}Z_{t}$

$\widehat{Y_{t}} = `r model1$coefficients[1]` + `r model1$coefficients[2]` X_{t} + `r model1$coefficients[3]` Z_{t}$

####Variables all together

```{r ols123, echo=FALSE, warning=FALSE, message=FALSE}
plot.zoo(cbind(data$Y, data$X, data$Z), 
                  plot.type = "single", 
                  col = c("red", "blue", "green"))
```

Variables Y, X and Z respectively colors red, blue and green.

#### Y
```{r ols12, echo=FALSE, warning=FALSE, message=FALSE}
plot(data$Y)
```

Therefore, we observe that there is an increasing trend in each variables. They do not well behave.

We obtain: $\widehat{Y_{t}} = \widehat{\beta_0} + \widehat{\beta_1}X_{t} + \widehat{\beta_2}Z_{t}$

$\widehat{Y_{t}} = `r model1$coefficients[1]` + `r model1$coefficients[2]` X_{t} + `r model1$coefficients[3]` Z_{t}$

#### X

```{r ols13, echo=FALSE, warning=FALSE, message=FALSE}
plot(data$X)
```

Therefore, we observe that there is an increasing trend in each variables. They do not well behave.

We obtain: $\widehat{Y_{t}} = \widehat{\beta_0} + \widehat{\beta_1}X_{t} + \widehat{\beta_2}Z_{t}$

$\widehat{Y_{t}} = `r model1$coefficients[1]` + `r model1$coefficients[2]` X_{t} + `r model1$coefficients[3]` Z_{t}$

#### Z

```{r ols14, echo=FALSE, warning=FALSE, message=FALSE}
plot(data$Z)
```

Therefore, we observe that there is an increasing trend in each variables. They do not well behave. 

We obtain: $\widehat{Y_{t}} = \widehat{\beta_0} + \widehat{\beta_1}X_{t} + \widehat{\beta_2}Z_{t}$

$\widehat{Y_{t}} = `r model1$coefficients[1]` + `r model1$coefficients[2]` X_{t} + `r model1$coefficients[3]` Z_{t}$


#### extra: Residuals for $Y_t = \beta_0 + \beta_1{X_t} + u_t$

Residuals
```{r olsa, echo=FALSE, warning=FALSE, message=FALSE}
resid(fit)
```

```{r ols556, echo=FALSE, warning=FALSE, message=FALSE}
plot(fit, which=1, col=c("blue"))
```

```{r olsq, echo=FALSE, warning=FALSE, message=FALSE}
ggplot(data, aes(x = X, y = Y)) +
  geom_smooth(method = "lm", se = FALSE, color = "lightgrey") +     # regression line  
  geom_segment(aes(xend = X, yend = predicted), alpha = .2) +      # draw line from point to line
  geom_point(aes(color = abs(residuals), size = abs(residuals))) +  # size of the points
  scale_color_continuous(low = "green", high = "red") +             # colour of the points mapped to residual size - green smaller, red larger
  guides(color = FALSE, size = FALSE) +                             # Size legend removed
  geom_point(aes(y = predicted), shape = 1) +
  theme_bw()
```

$\widehat{Y_{t}} = `r fit$coefficients[1]` + `r fit$coefficients[2]` X_{t}$

$\widehat{Y_{t}} = \widehat{\beta_0} + \widehat{\beta_1}X_{t} + \widehat{\beta_2}Z_{t}$

#### extra: Residuals for $Y_t = \beta_0 + \beta_1{Z_t} + u_t$

Residuals
```{r olsk, echo=FALSE, warning=FALSE, message=FALSE}
resid(fit2)
```

```{r ols5567, echo=FALSE, warning=FALSE, message=FALSE}
plot(fit2, which=1, col=c("blue"))
```

```{r olb7, echo=FALSE, warning=FALSE, message=FALSE}

ggplot(data, aes(x = Z, y = Y)) +
  geom_smooth(method = "lm", se = FALSE, color = "lightgrey") +     # regression line  
  geom_segment(aes(xend = Z, yend = predicted), alpha = .2) +      # draw line from point to line
  geom_point(aes(color = abs(residuals), size = abs(residuals))) +  # size of the points
  scale_color_continuous(low = "green", high = "red") +             # colour of the points mapped to residual size - green smaller, red larger
  guides(color = FALSE, size = FALSE) +                             # Size legend removed
  geom_point(aes(y = predicted), shape = 1) +
  theme_bw()
```


$\widehat{Y_{t}} = `r fit2$coefficients[1]` + `r fit2$coefficients[2]` Z_{t}$


#### 3D scatter plot

Residuals in 3D

```{r o49, echo=FALSE, warning=FALSE, message=FALSE}
library(rgl)
library(rglwidget)

mfrow3d(nr = 1, nc = 1, sharedMouse = T)  
scatter3d(Y ~ X + Z, fill=FALSE, data=data)
rglwidget()
```

### Durbin Watson Test for Autocorrelation

The estimation results of the following specification of the static model is as follows:
  
```{r ols3, echo=FALSE, warning=FALSE, message=FALSE}
db <- durbinWatsonTest(model1)
db
```
We observe positive DW value, it may have AC problem since we do not know the intervals of upper and lower d statistics. 
Prob[DW] = `r db$p` > 0.05 do not Rho. NO AC problem.

```{r ols49, echo=FALSE, warning=FALSE, message=FALSE}
dwtest(model1)
```

However, in other DW Test, our p value results 0.06. 

#### Breusch-Pagan Test for Autocorrelation
 
The estimation results of the following specification of the static model is as follows:

```{r ols4, echo=FALSE, warning=FALSE, message=FALSE}
bptest(model1)
```
Do not RHo, there is no HC at $\alpha$ = 0.05 level of significance.

# Task 2

### OLS Estimation of dynamic model
Task 2 requires the estimation of dynamic model, $Y_t = \alpha_0 + \alpha_1{Y_{t-1}} + \alpha_2{X_{t}} + \alpha_3{Z_{t}} + \alpha_4{X_{t-1}} + \alpha_5{Z_{t-1}} + \epsilon_t$ with ARDL(1,1,1).

```{r ols5, echo=FALSE, warning=FALSE, message=FALSE}

model2 <- lm(Y ~ X + Z + Ylag + Xlag + Zlag, data = data)
summary(model2)
```

Therefore, our model becomes 
$\widehat{Y_{t}} = 
`r model2$coefficients[1]` + 
`r model2$coefficients[2]` X_{t} + 
`r model2$coefficients[3]` Z_{t} + 
`r model2$coefficients[4]` Y_{t-1} + 
`r model2$coefficients[5]` X_{t-1} + 
`r model2$coefficients[6]` Z_{t-1}$

### COMFAC Test

